% really wanna talk about marginal gains when I first read itn James Clear's book, cite that here - https://jamesclear.com/marginal-gains
\epigraph{Success is the product of daily habits--not once-in-a-lifetime transformations}
{\textit{James Clear}}


\section{Marginal Gains}

In elite sport, there is an idea referred to as \textbf{marginal gains}\footnote{I first read about this in James Clear's Atomic Habits, an excerpt from which can be found \textbf{\href{https://jamesclear.com/marginal-gains}{\textit{here}}}}. The principle is simple: instead of looking for one large breakthrough, you focus on making many small improvements. Each small change on its own might seem insignificant, but together, they compound. 

One of the most well-known examples of this comes from British cycling. In the years leading up to their Olympic dominance, the team became almost absurdly committed to improving everything by tiny amounts. They monitored training, nutrition, recovery, sleep, equipment, even the things that seemed too small to be worth bothering with. Now what stood out was not the scale of any single improvement, but the accumulation of many small ones. 

\section{What Progress Usually Looks Like} 

In most settings, progress is often visible. Think about solving an equation step by step, with each step bringing you closer to a solution. When answering an exam question, you move through a sequence of known methods, gradually building towards a final answer. Even when the work is difficult, there is usually a clear sense of direction. You can see yourself getting closer.

This familiarity shapes how you expect progress to feel, and when those signals are absent, it is natural to assume that nothing is happening. 

At the beginning of a cipher, there is often no clear indication that you are getting anywhere. You may spend several minutes observing, testing ideas, and ruling things out, without producing anything that resembles a solution. Compared to more familiar problems, this can feel like a lack of progress.

\section{Why It Doesn’t Feel Like Progress} 

If this process is genuinely productive, why does it often feel like nothing is happening?

The difficulty is that small improvements do not feel like progress. There is no intermediate output to confirm that you are moving in the right direction. And without that feedback, it is easy to interpret the absence of a solution as the absence of progress.

If progress is defined only has having solved the cipher, then everything before that point starts to feel like failure. Which is a terrible definition, not least because it means you can spend hours doing genuinely intelligent work and still come away feeling like you have achieved nothing. This is the same mistake as ignoring marginal gains in sport because each individual improvement seems too small to matter. 

And I say that not from a position of great wisdom, but as someone who has absolutely stared at a ciphertext for ages, made several useful observations, and then convinced myself I was getting nowhere because none of them had yet turned into plaintext.

A lot of early insights only become meaningful in hindsight. A pattern that looks irrelevant at first may later turn out to be the thing everything hinged on. But in the moment, it rarely feels like revelation.


\section{Narrowing the Problem} 

At the start of a cipher, there are usually many possible interpretations. The text could represent letters, numbers, symbols, or groups. It could be the result of a simple substitution, a transposition, or one of those ciphers that makes you briefly question whether language itself was a mistake. Without additional information, these possibilities can feel equally plausible.

When you notice that certain patterns repeat in a consistent way, you begin to rule out some explanations. When the number of distinct symbols is too high or too low for a particular method, that method becomes less likely. When a hypothesis fails to produce consistent results, it can be set aside.

Over time, this process narrows the range of possibilities. You are not moving directly towards a solution, but making the problem smaller and more defined. Each step removes options until what remains is manageable enough to analyse more directly.

Looking at it this way, codebreaking is less about finding the answer and more about reducing what the answer could be.


\section{A More Useful Definition} 

Instead of asking: \textit{Have I solved the problem? }Try asking: \textit{Do I understand the problem better than I did before?}

If the answer is yes even in the slightest, then progress has been made. You may not yet have a solution, or may not even be close to one. But if you have ruled out possibilities, identified structure, or clarified your assumptions, then you have moved forward.

This shift in perspective is important because it changes how you approach the process. It allows you to value the marginal gains and the incremental steps that lead to understanding rather than focusing only on the final result. 
